---
title: "Reverse Engineering Project"
author: "Aline Behar Kado"
date: "October 2024"
output:
  html_document:
    theme: cerulean
    highlight: pygments
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

In this notebook, we are reverse engineering the story, [STORY NAME](STORY%20URL)

## Load libraries

Loading required libraries for this analysis.

```{r echo=FALSE, message=FALSE}
library(tidyverse)
library(lubridate)
library(janitor)
library(dplyr)

```

## Load and Cleaning Data

In this section, describe the source of the data, write a basic data dictionary for data you are working with, and discuss any caveats or issues you discovered working with this data.

```{r}
# Load required data
 pop_by_age <-read_csv("data/population-by-age.csv")
pub_enroll_by_grade <- read_csv("data/public-enrollment-by-grade.csv")
priv_enroll_by_grade <- read_csv("data/private-enrollment-by-grade.csv")
summary <- read_csv("data/Missing-Kids-Summary-File.csv")|>clean_names()

# Path to data should be loaded from folder "data" i.e. read_csv("data/name_of_data.csv")

# Clean required data and prepare for analysis if needed. 


```

## Sentences to Engineer

In this notebook, we are reverse engineering five sentences from the story.

### Sentence 1

-   **Sentence text**: Those states saw private-school enrollment grow by over 100,000 students. Home-schooling grew even more, surging by more than 180,000.
-   **Analysis summary**:

```{r}
# Put code to reverse engineer sentence here
states_used_priv_enroll <- priv_enroll_by_grade |> 
  filter(grade == "k_12_total")|>
   mutate(
    used = case_when(
        str_detect(state, "CA") ~ "True",
        str_detect(state, "NY") ~ "True",
        str_detect(state, "LA") ~ "True",
        str_detect(state, "NC") ~ "True",
        str_detect(state, "WA") ~ "True",
        str_detect(state, "CO") ~ "True",
        str_detect(state, "GA") ~ "True",
        str_detect(state, "MN") ~ "True",
        str_detect(state, "WI") ~ "True",
        str_detect(state, "NH") ~ "True",
        str_detect(state, "MA") ~ "True",
        str_detect(state, "MT") ~ "True",
        str_detect(state, "DE") ~ "True",
        str_detect(state, "ND") ~ "True",
        str_detect(state, "DC") ~ "True",
        str_detect(state, "SD") ~ "True",
        str_detect(state, "SC") ~ "True",
        str_detect(state, "NE") ~ "True",
        str_detect(state, "RI") ~ "True",
        str_detect(state, "TN") ~ "True",
        str_detect(state, "PA") ~ "True",
        str_detect(state, "FL") ~ "True",
        .default = "FALSE"
      )
  )
  
  
  

states_used_priv_enroll |> 
  filter(used == "True") |>
  group_by(sch_year)|>
  summarize(total = sum(enrollment, na.rm = TRUE)) |>
  arrange (desc(total))



# Display results of code below this codeblock

```

### Sentence 2

-   **Sentence text**: Overall, public school enrollment fell by over 700,000 students between the 2019-2020 and 2021-2022 school years in the 21 states plus Washington, D.C., that provided the necessary data.
-   **Analysis summary**:

```{r}
# Put code to reverse engineer sentence here

states_used_pub_enroll <- pub_enroll_by_grade |> 
  filter(grade == "k_12_total")|>
   mutate(
    used = case_when(
        str_detect(st, "CA") ~ "True",
        str_detect(st, "NY") ~ "True",
        str_detect(st, "LA") ~ "True",
        str_detect(st, "NC") ~ "True",
        str_detect(st, "WA") ~ "True",
        str_detect(st, "CO") ~ "True",
        str_detect(st, "GA") ~ "True",
        str_detect(st, "MN") ~ "True",
        str_detect(st, "WI") ~ "True",
        str_detect(st, "NH") ~ "True",
        str_detect(st, "MA") ~ "True",
        str_detect(st, "MT") ~ "True",
        str_detect(st, "DE") ~ "True",
        str_detect(st, "ND") ~ "True",
        str_detect(st, "DC") ~ "True",
        str_detect(st, "SD") ~ "True",
        str_detect(st, "SC") ~ "True",
        str_detect(st, "NE") ~ "True",
        str_detect(st, "RI") ~ "True",
        str_detect(st, "TN") ~ "True",
        str_detect(st, "PA") ~ "True",
        str_detect(st, "FL") ~ "True",
        .default = "FALSE"
      )
  )
  
  
  

states_used_pub_enroll |> 
  filter(used == "True") |>
  group_by(sch_year)|>
  summarize(total = sum(enrollment, na.rm = TRUE)) |>
  arrange (desc(total))




# Display results of code below this codeblock

```

### Sentence 3

-   **Sentence text**: The number of children registered for home-schooling surged by **around 184,000**.
-   **Analysis summary**: [Write up two to three sentences describing the results of your analysis. Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here

homeschool <- summary|>
  filter(enrollment_type == "Homeschool enrollment")|>
  mutate(
    used = case_when(
        str_detect(state_abbrev, "CA") ~ "True",
        str_detect(state_abbrev, "NY") ~ "True",
        str_detect(state_abbrev, "LA") ~ "True",
        str_detect(state_abbrev, "NC") ~ "True",
        str_detect(state_abbrev, "WA") ~ "True",
        str_detect(state_abbrev, "CO") ~ "True",
        str_detect(state_abbrev, "GA") ~ "True",
        str_detect(state_abbrev, "MN") ~ "True",
        str_detect(state_abbrev, "WI") ~ "True",
        str_detect(state_abbrev, "NH") ~ "True",
        str_detect(state_abbrev, "MA") ~ "True",
        str_detect(state_abbrev, "MT") ~ "True",
        str_detect(state_abbrev, "DE") ~ "True",
        str_detect(state_abbrev, "ND") ~ "True",
        str_detect(state_abbrev, "DC") ~ "True",
        str_detect(state_abbrev, "SD") ~ "True",
        str_detect(state_abbrev, "SC") ~ "True",
        str_detect(state_abbrev, "NE") ~ "True",
        str_detect(state_abbrev, "RI") ~ "True",
        str_detect(state_abbrev, "TN") ~ "True",
        str_detect(state_abbrev, "PA") ~ "True",
        str_detect(state_abbrev, "FL") ~ "True",
        .default = "FALSE"
      )
  )

homeschool|>
  filter(used == "True") |>
  summarize(total = sum(total_change, na.rm = TRUE)) |>
  arrange (desc(total))

# Display results of code below this codeblock

```

### Sentence 4

-   **Sentence text**: [Paste in sentence to engineer here]
-   **Analysis summary**: [Write up two to three sentences describing the results of your analysis. Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here

# Display results of code below this codeblock

```

### Sentence 5

-   **Sentence text**: [Paste in sentence to engineer here]
-   **Analysis summary**: [Write up two to three sentences describing the results of your analysis. Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here

# Display results of code below this codeblock

```

-30-
